---
title: "Data_Challenge_3_AAnderson"
subtitle: "GitHub Link: "
author: "Ashley Anderson"
date: "10/28/2020"
output: html_document
---
```{r,echo=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv') ## read in the data
library(dplyr) ## load libraries
library(ggplot2)
library(tidyr)
library(stringr)
library(purrr)
```

**Question 1: Write a function that returns a table with the mean and standard deviation of the aroma through moisture columns of the coffee ratings data for different categories of coffee. Produce the table for (a) species, (b) country of origin, and (c) binned mean altitude in meters.**

```{r, eval = FALSE}
## Decide how to bin mean altitude
coffee_ratings_mean <- filter_at(coffee_ratings, vars(altitude_mean_meters), any_vars(. < 10000)) ## Looking at the data, there were 4 extreme values greater than 10,000 meters. Those will be put in their own bin.
median(coffee_ratings_mean$altitude_mean_meters) ## Calculating the median of the data decided the cutoff for the other two bins = 1311
```
```{r}
coffee_ratings_binned <- coffee_ratings %>%
  mutate(binned_mean_altitude = ## rename new column for binned altitude
           cut(altitude_mean_meters, ## cut from the altitude_mean_meters column
               breaks = c(0, 1311, 10000, Inf), ## bin by low and high by median and then extreme values greater than 10,000 meters
               labels = c("low (0 to 1310.9 meters)", "high (1311 to 10,000 meters)", "extreme (10,000 meters+)"))) ## add labels for each bin

ratings_1 <- function(x) { ## initiate function with one variable, x
  ratings <- paste(round(mean(x), 2), ## paste the mean of the variable rounded to 2 decimals
                   paste0("[sd: ", round(sd(x), 2), "]")) ## paste the standard deviation of the variable rounded to 2 decimals with square brackets and descriptive text
  return(ratings) ## return data
}

ratings_table_fxn <- function(df, x) { ## initiate function with a data frame (df) and one variable, x
  ratings_table <- df %>%
    rename(group = x) %>% ## rename column referred to by string 'x' as group
    group_by(group) %>% ## group by "group"
    summarize(across(aroma:moisture, ratings_1)) %>% ## summarize data across numerical value columns of data frame; pass through the mean and sd function to calculate statistics for each column
    knitr::kable() ## initiate kable table
  return(ratings_table) ## return data
}

Species_ratings <- ratings_table_fxn(coffee_ratings_binned, "species")
Country_ratings <- ratings_table_fxn(coffee_ratings_binned, "country_of_origin")
Altitude_ratings <- ratings_table_fxn(coffee_ratings_binned, "binned_mean_altitude")
Species_ratings
Country_ratings
Altitude_ratings
```

**Question 2: Write a function that makes a scatterplot of any two variables from the columns aroma:moisture and colors by a particular grouping. Make the plot for (1) aroma versus flavor colored by species and (2) balance versus body colored by binned mean altitude.**

```{r}
coffee_ratings_filter1 <- filter_at(coffee_ratings_binned, vars(aroma:balance), any_vars(. > 0)) ## removed "outliers" of 0 from aroma:balance columns
coffee_ratings_filter2 <- filter_at(coffee_ratings_filter1, vars(body), any_vars(. > 5.99)) ## removed handful of "outliers" from body column

ratings_plot <- function(df, x, y, colorby) { ## setting up function to pass the data frame and 3 variables
  df %>% ## pipe in data frame
    ggplot(aes(x = {{x}}, y = {{y}}, col = {{colorby}})) + ## initiate ggplot with x, y and the "color by" variable
    geom_point() + ## create scatterplot
    labs(title = paste("coffee", as_label(enquo(x)), "versus", as_label(enquo(y)), "colored by", as_label(enquo(colorby)))) ## add title 
}

Plot1 <- ratings_plot(coffee_ratings_filter2, aroma, flavor, species) ## Plot aroma versus flavor colored by species using "ratings_plot" function created above
Plot2 <- ratings_plot(coffee_ratings_filter2, balance, body, binned_mean_altitude) ## Plot balance versus body colored by binned mean altitude using "ratings_plot" function created above
Plot1
Plot2
```



